Incentivizing Exploration by Heterogeneous Users

### 2018-02-14 lower_bound.tex
did some minor changes to the paper
lower bound changed to square

We saw in Corollary~\ref{cor:constant-regret} that
when agent preferences for their best arm are sufficiently clear,
in the sense that $\TieDensity = 0$, the regret of
Algorithm~\ref{alg:basic-incentivizing} is bounded by a constant.
One may conjecture that this should hold more generally,
in that that regret of the (fewer and fewer) agents on the boundary
between close arms should go to 0, while their fraction also goes to 0.
In this section, we establish a lower bound,
showing that even in very simple settings, a (logarithmic) dependence
on $T$ is typicaly unavoidable for \emph{any} incentivization strategy.

We consider an instance with two arms,
whose attribute vectors are $(0,0)$ and $(0,1)$, respectively.
Agents types are distributed uniformly on the two-dimensional unit circle%
\footnote{While our model technically assumed that all agent coordinates are
non-negative, we could simply shift the unit circle.
The present choice is solely for ease of notation.}  
$\Set{\AgV}{|| \AgV ||_2 \leq 1}$.
Because $\AgV \cdot (0,0) = 0$ and $\AgV \cdot (0,1) = \Ag{2}$,
the best choice for agent \AgV is arm $(0,1)$ iff $\Ag{2} > 0$,
i.e., the top half of the unit circle prefers the arm $(0,1)$,
and the bottom half prefers the arm $(0,0)$.

@@

For the moment, focus on some time step $t$,
and write $\NoiseV = \NoiseV[t]$.
Then, there are two types of agents who pull the wrong arm and incur
regret:
\begin{enumerate}
\item If $\Ag{2} > 0$ and $\Ag{1} \Noise{1} + \Ag{2} (1+\Noise{2}) < 0$
then \AgV should pull $(0,1)$,
but will wrongly pull $(0,0)$ and incur regret \Ag{2}.
The region
$\Ag{2} > 0$ and $\Ag{1} < \frac{-(1+\Noise{2})}{\Noise{1}} \cdot \Ag{2}$
defines a wedge.
For any particular value of \Ag{2},
the length of the interval of corresponding \Ag{1} is
\begin{align*}
  \sqrt{1-\Ag{2}^2} - \frac{1+\Noise{2}}{\Noise{1}} \cdot \Ag{2}
  & \geq 1 - \Ag{2} - \frac{1+\Noise{2}}{\Noise{1}} \cdot \Ag{2}
  \; = \; 1 - \frac{1+\Noise{1}+\Noise{2}}{\Noise{1}} \cdot \Ag{2}.
\end{align*}

\item If $\Ag{2} < 0$ and $\Ag{1} \Noise{1} + \Ag{2} (1+\Noise{2}) > 0$,
then \Ag should pull $(0,0)$,
but will wrongly pull $(0,1)$ and incur regret $-\Ag{2}$.
This region and its regret are rotationally symmetric to the previous
case.
\end{enumerate}

Hence, the expected regret for given \Noise{1}, \Noise{2} is at least
\begin{align*}
  2 \int_0^{\frac{\Noise{1}}{1+\Noise{1}+\Noise{2}}}
  \left( 1 - \frac{1+\Noise{1}+\Noise{2}}{\Noise{1}} \cdot \Ag{2} \right)
  \cdot \Ag{2} \cdot \frac{1}{\pi} \dd \Ag{2}
& =
\frac{1}{\pi} \cdot \left(\frac{\Noise{1}}{1+\Noise{1}+\Noise{2}}\right)^2
- \frac{2}{3 \pi} \cdot \left(\frac{\Noise{1}}{1+\Noise{1}+\Noise{2}}\right)^2
\\ & =
\frac{1}{3 \pi} \cdot \left(\frac{\Noise{1}}{1+\Noise{1}+\Noise{2}}\right)^2.
\end{align*}

\begin{align}
&E[r(t)| \Noise{1}>0, \Noise{2}>0] \nonumber \\
\geq & 4\times \frac{1}{3\pi} \int_{0}^{\infty} \int_{0}^{\infty} (\frac{\Noise{1}}{1+\Noise{1}+\Noise{2}})^2\frac{e^{-\frac{t \times \Noise{1}^2}{2}}\sqrt{t}}{\sqrt{2\pi}}d(\Noise{1})\frac{e^{-\frac{t \times \Noise{2}^2}{2}}\sqrt{t}}{\sqrt{2\pi}}d(\Noise{2}) \nonumber \\
= & \frac{1}{3\pi^3} \int_{0}^{\infty} \int_{0}^{\infty} (\frac{\Noise{1}}{\sqrt{t}+\Noise{1}+\Noise{2}})^2e^{-\frac{\Noise{1}^2}{2}}d(\Noise{1})e^{-\frac{\Noise{2}^2}{2}}d(\Noise{2}) \nonumber 
\end{align}

Below, we want to show 
\begin{align}
\lim_{t\rightarrow\infty}\frac{E[r(t)| \Noise{1}>0, \Noise{2}>0]}{\frac{1}{t}} = \Omega(1), \nonumber
\end{align}
and use the fact that $\sum_{n=1}^{T}\frac{1}{n}=O(\log(T))$ to show the regret is at least $\Omega(\log(T))$. Since
\begin{align}
&\lim_{t\rightarrow \infty}\frac{E[r(t)| \Noise{1}>0, \Noise{2}>0]}{\frac{1}{t}} \nonumber \\
\geq & \lim_{t\rightarrow \infty}\frac{1}{3\pi^3}\int_{0}^{\infty} \int_{0}^{\infty}\left[ t(\frac{\Noise{1}}{\sqrt{t}+\Noise{1}+\Noise{2}})^2\right]e^{-\frac{\Noise{1}^2}{2}}e^{-\frac{\Noise{2}^2}{2}}d(\Noise{1})d(\Noise{2}) \nonumber  \\
= & \frac{1}{3\pi^3}\int_{0}^{\infty} \int_{0}^{\infty}\lim_{t\rightarrow \infty}\left[ t(\frac{\Noise{1}}{\sqrt{t}+\Noise{1}+\Noise{2}})^2\right]e^{-\frac{\Noise{1}^2}{2}}e^{-\frac{\Noise{2}^2}{2}}d(\Noise{1})d(\Noise{2}) \nonumber  \\
=&\frac{1}{3\pi^3}\int_{0}^{\infty} \int_{0}^{\infty}e^{-\frac{\Noise{1}^2}{2}}e^{-\frac{\Noise{2}^2}{2}}d(\Noise{1})d(\Noise{2}) = \frac{1}{3\pi^3}, \nonumber
\end{align}

where the third line is due to monotone convergence theorem since the derivative of $t(\frac{\Noise{1}}{\sqrt{t}+\Noise{1}+\Noise{2}})^2$ is greater than $0$.

\dkcomment{I edited up to here. Waiting for some feedback about
  whether my calculations are way off.
  I did use one bound of $\sqrt{1-\Ag{2}^2} \geq 1-\Ag{2}$, but since
  we are looking at small values of \Ag{2}, I cannot imagine that this
  lost much. One can compute the integral without this simplification,
  but it will be quite a bit uglier.}

>>>>

We saw in Corollary~\ref{cor:constant-regret} that
when agent preferences for their best arm are sufficiently clear,
in the sense that $\TieDensity = 0$, the regret of
Algorithm~\ref{alg:basic-incentivizing} is bounded by a constant.
One may conjecture that this should hold more generally,
in that the regret of the (fewer and fewer) agents on the boundary
between close arms should go to 0, while their fraction also goes to 0.
In this section, we establish a lower bound,
showing that even in very simple settings, a (logarithmic) dependence
on $T$ is typicaly unavoidable for \emph{any} incentivization strategy.

We consider an instance with two arms,
whose attribute vectors are $(0,0)$ and $(0,1)$, respectively.
Agents types are distributed uniformly on the (edge of the)
two-dimensional unit square%
\footnote{While our model technically assumed that all agent coordinates are
non-negative, we could simply shift the unit square.
The present choice is solely for ease of notation.}  
$\Set{\AgV}{\max(\Ag{1}, \Ag{2}) = 1}$,
with density $\frac{1}{8}$.
Because $\AgV \cdot (0,0) = 0$ and $\AgV \cdot (0,1) = \Ag{2}$,
the best choice for agent \AgV is arm $(0,1)$ iff $\Ag{2} > 0$,
i.e., the top half of the unit circle square the arm $(0,1)$,
and the bottom half prefers the arm $(0,0)$.

@@

For the moment, focus on some time step $t$,
and write $\NoiseV = \NoiseV[t]$.
Then, there are two types of agents who pull the wrong arm and incur
regret:
\begin{enumerate}
\item If $\Ag{2} > 0$ and $\Ag{1} \Noise{1} + \Ag{2} (1+\Noise{2}) < 0$
then \AgV should pull $(0,1)$,
but will wrongly pull $(0,0)$ and incur regret \Ag{2}.
The range of \AgV making the wrong choice is thus
$0 < \Ag{2} < \frac{- \Noise{1}}{1+\Noise{2}} \cdot \Ag{1}$.
Since we are proving a lower bound, we only focus on the
case $\Ag{1} = -1$, and ignore the case $\Ag{2} = 1$
(which is rare, since it requires \Noise{1} to be large).
Thus, the set of agents incurring regret contains the set
$\Set{\AgV = (-1, \Ag{2})}{0 < \Ag{2} < \frac{\Noise{1}}{1+\Noise{2}}}$.

% For any particular value of \Ag{2},
% the length of the interval of corresponding \Ag{1} is
% \begin{align*}
%   \sqrt{1-\Ag{2}^2} - \frac{1+\Noise{2}}{\Noise{1}} \cdot \Ag{2}
%   & \geq 1 - \Ag{2} - \frac{1+\Noise{2}}{\Noise{1}} \cdot \Ag{2}
%   \; = \; 1 - \frac{1+\Noise{1}+\Noise{2}}{\Noise{1}} \cdot \Ag{2}.
% \end{align*}

\item If $\Ag{2} < 0$ and $\Ag{1} \Noise{1} + \Ag{2} (1+\Noise{2}) > 0$,
then \AgV should pull $(0,0)$,
but will wrongly pull $(0,1)$ and incur regret $-\Ag{2}$.
This region and its regret are rotationally symmetric to the previous
case.
\end{enumerate}

Hence, the expected regret for given \Noise{1}, \Noise{2} is at least
%\begin{align*}
$2 \int_0^{\frac{\Noise{1}}{1+\Noise{2}}}
  \Ag{2} \cdot \frac{1}{8} \dd \Ag{2}
=
\frac{1}{8} \cdot \left(\frac{\Noise{1}}{1+\Noise{2}}\right)^2$.
%\end{align*}
The expected regret, conditioned on
$\Noise{1} > 0$ and $\Noise{2} > 0$, is therefore

\dkcomment{Where did the 4 come from? Where did it go? Also, why did
  you have $\pi^3$ (instead of $\pi^2$) before? (I changed it.)}
\begin{align*}
\ExpectC{\Regret{t}}{\Noise{1} > 0, \Noise{2} > 0}
 & \geq \dkedit{4??} \cdot \frac{1}{8}
    \int_{0}^{\infty} \int_{0}^{\infty}
    \left( \frac{\Noise{1}}{1+\Noise{2}} \right)^2 \cdot
    \frac{\e^{-\frac{t \Noise{1}^2}{2}} \sqrt{t}}{\sqrt{2\pi}} \dd\Noise{1}
    \frac{\e^{-\frac{t \Noise{2}^2}{2}} \sqrt{t}}{\sqrt{2\pi}} \dd\Noise{2} \\
 & = \frac{1}{16\pi} \int_{0}^{\infty} \int_{0}^{\infty}
   \left( \frac{\Noise{1}}{\sqrt{t}+\Noise{2}} \right)^2
   \e^{-\Noise{1}^2/2} \e^{-\Noise{2}^2/2} \dd \Noise{1} \dd \Noise{2}.
\end{align*}

We want to show that the expected regret per time step decreases only
at a rate of $\Omega(1/t)$, and thereto consider the limit of the
ratio of the two quantities:

\begin{align*}
\lim_{t \to \infty} \frac{\ExpectC{\Regret{t}}{\Noise{1} > 0, \Noise{2} > 0}}{\frac{1}{t}}
  & \geq
    \lim_{t \to \infty} \left( \frac{1}{16\pi} \cdot
    \int_{0}^{\infty} \int_{0}^{\infty}
    t \cdot \left( \frac{\Noise{1}}{\sqrt{t}+\Noise{2}} \right)^2
    \e^{-\Noise{1}^2/2} \e^{-\Noise{2}^2/2} \dd \Noise{1} \dd \Noise{2} \right) \\
  & = \frac{1}{16\pi} \cdot \int_{0}^{\infty} \int_{0}^{\infty}
    \lim_{t \to \infty}
    \left( t \cdot \left( \frac{\Noise{1}}{\sqrt{t}+\Noise{2}} \right)^2 \right)
    \e^{-\Noise{1}^2/2} \e^{-\Noise{2}^2/2} \dd \Noise{1} \dd \Noise{2} \\
  & = \frac{1}{16\pi} \cdot \int_{0}^{\infty} \int_{0}^{\infty}
    \e^{-\Noise{1}^2/2} \e^{-\Noise{2}^2/2} \dd \Noise{1} \dd \Noise{2}
  \; = \; \frac{1}{16 \pi}; 
\end{align*}

the second line is due to the monotone convergence theorem,
which can be applied because
$t \left(\frac{\Noise{1}}{\sqrt{t}+\Noise{2}}\right)^2$
is strictly increasing in $t$.
Because the expected regret in each time step is at least
$\Omega(1/t)$, the total expected regret is at least
$\Omega(\sum_{t=1}^{T}\frac{1}{t} = \Omega(\log(T))$.
